Network Configuration Synthesis with Abstract Topologies

### 2016-11-11 paper/abstraction/paper.tex
edits
working through section 4

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}
\usepackage{url}
\usepackage{times}
\usepackage{xspace}
\usepackage{listings}
\usepackage{code}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{tikz}
\usetikzlibrary{arrows,automata,positioning}
\let\proof\relax
\let\endproof\relax
\usepackage{amsthm}
\usepackage{subcaption}
\usepackage{balance}
\usepackage[ampersand]{easylist}
\usepackage{textcomp}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

@@

To summarize, the central contributions of this paper are:

\begin{enumerate}
\item Design of new topology abstractions for network programming
based on graph homomorphisms and connectivity invariants.
\item Design of two new algorithms for analyzing routing policy over
abstract networks: one which
guarantees correct compilation and one which determines fault tolerance
properties of network policy.  Both algorithms provide guarantees for any
concrete network satisfying the abstract connectivity invariants.
\item Design of template generation algorithms that support network
evolution.
\item Implementation and evaluation of a \sysname compiler that generates
configuration templates and is up to two orders of magnitude faster
than its closest competitor on large networks.
\end{enumerate}

@@

\sysname is a system for generating provably-correct templates from the network's policy and an {\em abstract} topology that describes device roles and connectivity invariants.
%As inputs, it takes a global, network-wide policy (as opposed to a per-device or even per-role policy) and an {\em abstract} topology that describes device roles and connectivity invariants.  It outputs a template for each role, from which per-device configurations can be generated by instantiating parameters with values drawn from the network topology.
It guarantees that:
(1) the configurations generated from its templates correctly implement the network's policy for {\em any} concrete topology that matches the abstract topology; and (2) if nodes or edges are added or removed in a way compatible
with the abstract topology, only configurations of adjacent devices ever need to be updated.

This section introduces \sysname using the examples in \S\ref{sec:motivation}, and the following sections detail template generation and formally define the correctness guarantees.

%As we have seen, writing correct network configurations is challenging task. Operators must manually decompose network-wide policies into distributed device configurations, reason about all combinations of network failures, and parameterize configurations to allow for change, ideally in a way that is amenable to incremental change. In this section, we give an overview of the \sysname system, which gives operators a principled way to accomplish all of these tasks.

%\subsection{Concrete Methane}%

%Consider again the datacenter from figure~\ref{fig:example}. To configure the datacenter routing
%policy, we would start by writing the following definition in \sysname:
%%
%\begin{code}
%\Define Routing =
%    \{P1 \Path \End(T1)
%     P2 \Path \End(T2)
%     P3 \Path \End(T3)
%     P4 \Path \End(T4)
%     P5 \Path \End(T5)
%     P6 \Path \End(T6)
%     P7 \Path \End(T7)
%     P8 \Path \End(T8)
%     \True \Path \Exit(Peer1 \Prefer Peer2) \}
%\end{code}
%\noindent%%

%This defines a group of constraints for basic routing behavior. The first line states
%that traffic destined for prefix $P1$ must follow a path that ends up at destination router $T1$.
%A similar constraint is added for each of the prefixes P[2-8]. The final rule matches all other
%prefix destinations and allows traffic to follow a path that leaves the datacenter through either
%of its peers (Peer1 or Peer2) with a preference for leaving through Peer1. The \Prefer symbol indicates
%that traffic should satisfy the constraint on the left whenever possible, and only resort to the backup (right)
%when this is not possible due to failures.%

%Next we can add the locality constraint for prefixes P[5-8]. We do this by adding another group of constraints:
%%
%\begin{code}
%\Define Locality =
%    \{P5 | P6 | P7 | P8 \Path \Always(\In)\}
%\end{code}
%\noindent%%

%The Locality constraint applies to any traffic destined for a local prefix and adds
%the constraint that the traffic must follow a path that matches an internal location
%(\IE, a router inside the datacenter) at each hop of the path. This prevents traffic
%from ever entering from or leaving through an external peer.%

%Finally, for all traffic in the datacenter, we will disallow valley routing. This constraint
%can be captured with the following additions to the policy:
%%
%\begin{code}
%\Define Tor = \{T1,T2,T3,T4,T5,T6,T7,T8\}
%\Define Agg = \{A1,A2,A3,A4,A5,A6,A7,A8\}
%\Define Spn = \{S1,S2\}
%\Define ValleyFree =
%    \{\True \Path \Novalley(Tor,Agg,Spn)\}
%\end{code}
%\noindent%%

%The first three lines define groups of routers. The top of rack (Tor) routers T[1-8], the
%aggregation routers (Agg) A[1-8], and the spine routers (Spn) S[1-2]. The next line
%adds a new constraint that applies to all traffic and ensures that traffic never follows a
%path that forms a valley between any of these three groups (\EG, going from Spn to Agg to Spn again).
%As we will see in section \ryan{reference}, each of these constraints such as the valley free constraint
%are really just syntactic sugar for regular expressions over topology locations.%

%The final policy is declared to as the conjunction of each of these constraints:
%%
%\begin{code}
%\Define Main =
%    Routing &
%    Locality &
%    ValleyFree &
%    \Agg(PG, \In -> \Out)
%\end{code}
%\noindent%
%The last part of the policy declares that we want to perform aggregation for global prefixes
%by summarizing them into the more general prefix PG at the border of the datacenter
%(\IE along any edge starting inside our datacenter and ending outside).%

%From this policy, \sysname can synthesize device-level BGP configurations for every router in
%the network that guarantee correct forwarding behavior under all possible failure combinations.
%Synthesizing configurations for each device from a network-wide specification both simplifies
%network configuration by writing policy in one place rather than dozens and simultaneously prevents
%a large class of errors associated with manual configuration by construction. For example,
%the compiler ensures configurations are always in sync and prevents the operator from having
%to manually consider failure scenarios.

@@

\sysname takes as input the network's abstract topology and its policy. The policy has two components: the routing policy describes how traffic should flow, and the fault-tolerance policy describes how many simultaneous failures a the network can withstand without loosing connectivity.

\para{Abstract topology}
Abstract topologies in \sysname capture the basic connectivity structure of the network, which holds when the network is originally designed and as it evolves. They are encoded in the form of a graph homomorphism annotated with logical constraints about node and edge multiplicities. We designed these abstractions to be able to precisely capture most real network topologies, while being amenable to the fault tolerance analysis and configuration synthesis that is done by our compiler.

@@

We introduce our topology abstractions using the networks from \S\ref{sec:motivation}.
The first is a role-based abstraction that allows an operator to map routers in the concrete network to roles in the abstract network. Figure~\ref{fig:example3} shows an example of an abstraction for both datacenter networks from \S\ref{sec:motivation}. In particular the concrete networks are abstracted into a new topology with 5 different roles: local ToR (TL), global ToR (TG), local aggregator (AL), global aggregator (AG), and spine (S).

More specifically, a network topology is a graph $G$ = ($V, E$), which consists of a set of vertices $V$ and a set of directed edges $E \colon V \times V$. A role-based abstraction is a graph homomorphism from $G$ to an abstract graph $G^A$ = ($V^A$,$E^A$). A graph homorphism $f : G \rightarrow G^A$ is a function that maps each node in the concrete graph to a node in the abstract graph such that, whenever $(u,v) \in E$, then $(f(u),f(v)) \in E^A$.

@@

The role-based abstraction therefore overapproximates the connectivity of the underlying concrete graphs. Unfortunately, by itself, this abstraction loses a lot of information about the underlying structure of the concrete network, which would hinder a useful analysis of its fault-tolerance. For example, with this abstraction any spine router in a concrete network may or may not be connected to any aggregator router.
%
In order to refine the abstraction to capture more precisely the kinds of concrete networks of interest,
we introduce a notion of node and edge \emph{multiplicities}. Each edge (and node) is labelled
with a symbolic variable (\EG, $e1, e2, \ldots$) that denotes a constraint on the number of nodes/edges
that may appear in any given concrete network. Operators can then capture concrete network invariants by adding constraints on the symbolic variables using logical formula.

For example, in Figure~\ref{fig:example3}, the first two constraints: $e1 = AG$ and $e2 = TG$ declare that, within any pod P, the number of outgoing edges from $TG$ to $AG$ is equal to the number of nodes that map to $AG$. Similarly, the number of outgoing edges from $AG$ to $TG$ is equal to the number of nodes that map to $TG$. These constraints capture the fact that, within any given pod, the global aggregators and ToRs are in a full mesh. Furthermore, the constraints $TG = AG$ and $TG \geq 2$ ensure that the number of concrete nodes in role $TG$ is always equal to the number of concrete nodes in role $AG$, which is at least 2. Similar constraints also appear for the local aggregators and ToRs.
%
The constraint $e3 \geq 2$ says that, in each pod, each aggregator node has at least 2 outgoing edges to nodes in the spine role. Symmetrically, the constraint $e4 \geq 1$ says that, for each pod P, each node in the spine role has at least one outgoing edge to a node in the $AG$ role. Again, similar constraints appear for the local aggregator role. Finally, the constraints $S \geq 2$, $(S \mod AG) = 0$, and $(S \mod AL) = 0$ ensure that there are at least 2 routers in the spine role, and that the number of spine routers is always divisible by the number of global and local aggregators.
%
The last two constraints are interesting in that they do not involve a simple inequality. In general operators can write constraints using logical formula for any theory supported by modern SMT solvers.

@@

\para{Routing policy} Routing policies in \sysname consist of mappings from predicates matching various classes of traffic to a list of ranked constraints on the paths that the traffic matching the predicate should prefer to take through the network. We borrow syntax from Propane~\cite{propane} but, instead of concrete prefixes and nodes, \sysname specifications use abstract prefixes and nodes.

Let us see how to configure each of the datacenter networks from \S\ref{sec:motivation} by writing a \sysname routing policy over the new abstract topology. We can capture the basic routing behavior for the datacenters with the following constraints:
%
\begin{code}
\Define Routing =
    \$GP  \Path \End(TG)
    \$LP  \Path \End(TL)
    \True \Path \Exit(Peer1 \Prefer Peer2)
\end{code}
\noindent%

The first line introduces a prefix \emph{template} variable. It says that traffic for each global prefix associated with the template variable \$GP has the constraint that it should follow a path that ends at its corresponding destination router contained in the abstract TG role. The second line does the same thing for all local prefixes.
%
The final rule matches all other IP prefix destinations and allows traffic to follow a path that leaves the datacenter through either of its peers (Peer1 or Peer2) with a preference for leaving through Peer1. The \Prefer~symbol indicates that traffic should satisfy the constraint on the left whenever possible (\IE, a path exists in the network), and only resort to the backup (right) when this is not possible due to failures.%

Next we can capture the constraint that traffic for local prefixes must stay local to the datacenter.
We do this with the following \sysname constraint:
%
\begin{code}
\Define Locality =
    \$LP \Path \Always(\In)
\end{code}
\noindent%
%
The Locality constraint applies to any traffic destined for a local prefix described by the \$LP prefix template variable and adds the constraint that the traffic must follow a path that matches an internal location (\IE, a router inside the datacenter) at each hop of the path. This prevents traffic from ever entering from or leaving through an external peer. Finally, we can combine these constraints together as follows:

\begin{code}
Routing & Locality & \Agg(PG, \In -> \Out)
\end{code}
\noindent%

\ryan{we may need to add notransit constraint to make the compilation section simpler}

This tells \sysname to find BGP router configurations that satisfy the conjunction of the Routing and Locality constraints. It also declares an additional constraint that we want to perform prefix summarization (as PG) for global prefixes at the border of the datacenter (\IE, along any edge starting inside our datacenter and ending at a peer outside the datacenter).

When \sysname compiles the above policy, it guarantees that the forwarding behavior of the network will be correct for any concrete instantiation of the abstract network, and for any combination of failures that may occur in the concrete network. Furthermore, \sysname will generate configurations in such a way that, if we want to expand our network, for example by adding or removing a spine routers as in figure~\ref{fig:example2}, then the expansion can be performed by only ever touching device configurations that are adjacent to the change.

@@

%For example, the combining the Routing and Locality constraints in the data center policy from \S\ref{sec:overview} would result in the following \sysname policy:
%
%\begin{code}
%\$PG  \Path \End(TG),
%\$PL  \Path \End(TL) \ensuremath{\cap} \Always(\In),
%\True \Path \Exit(Peer1) \Prefer \Exit(Peer2)
%\end{code}
%\noindent%
%

@@

%By combining the tagging, filtering and preference mechanisms, we can generate safe BGP configurations. In general, configurations are generated by combining the per-prefix configurations in the same order specified in the policy.
Figure~\ref{fig:bgp-configs} shows part of the final configuration for spine routers for the concrete and abstract policies.
%
For prefix $0.0.0.0/[0..32]$, the spine routers will match advertisements from peer $P1$ and $P2$. The match for $P1$ is preferred since it has a higher BGP local-preference attribute (110). If an advertisement from $P1$ is available the spine uses this and attaches the community tag $(1,0)$ and advertises the route to all its peers. If only an advertisement from the backup $P2$ is available, then it attaches the tag $(0,1)$ and advertises this to all peers instead.
%
There is also additional policy for the Product Graphs from constraints for the local and global prefixes. In the template configuration, it matches any global prefix from any internal peer and re-advertise the route to all of its peers. For any local prefix it will allow an advertisement from any internal peer, and re-advertise the route to only other internal peers (thus implementing the Locality constraint). The concrete configurations have a similar configuration for each local and global prefix.

>>>>

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}
\usepackage{url}
\usepackage{times}
\usepackage{xspace}
\usepackage{listings}
\usepackage{code}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{tikz}
\usetikzlibrary{arrows,automata,positioning}
\let\proof\relax
\let\endproof\relax
\usepackage{amsthm}
\usepackage{subcaption}
\usepackage{balance}
\usepackage[ampersand]{easylist}
\usepackage{textcomp}
\usepackage{enumitem}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

@@

\vspace{1em}
\noindent
\textbf{Contributions}: To summarize, the contributions are:
%
\begin{itemize}[noitemsep,nolistsep,label={\large\textbullet}]
\item Design of new topology abstractions for network programming
based on graph homomorphisms and connectivity invariants.
\item Design of two new algorithms for analyzing routing policy over
abstract networks: one which
guarantees correct compilation and one which determines fault tolerance
properties of network policy.  Both algorithms provide guarantees for any
concrete network satisfying the abstract connectivity invariants.
\item Design of template generation algorithms that support network
evolution.
\item Implementation and evaluation of a \sysname compiler that generates
configuration templates and is up to two orders of magnitude faster
than its closest competitor on large networks.
\end{itemize}

@@

\sysname is a system for generating provably-correct templates from the network's policy and an {\em abstract} topology that describes device roles and connectivity invariants.
%As inputs, it takes a global, network-wide policy (as opposed to a per-device or even per-role policy) and an {\em abstract} topology that describes device roles and connectivity invariants.  It outputs a template for each role, from which per-device configurations can be generated by instantiating parameters with values drawn from the network topology.
%It guarantees that:
%(1) the configurations generated from its templates correctly implement the network's policy for {\em any} concrete topology that matches the abstract topology; and (2) if nodes or edges are added or removed in a way compatible
%with the abstract topology, only configurations of devices adjacent to the change ever need to be updated.
%
%This section introduces \sysname using the examples in \S\ref{sec:motivation}.
%, and the following sections detail template generation and formally define the correctness guarantees.
%
When \sysname compiles a policy, it guarantees that the forwarding behavior of the network will be correct for any concrete instantiation of the abstract network, and for any combination of failures that may occur in the concrete network. Furthermore, \sysname will generate configurations in such a way that, if an operator expands their network in a way compatible with the abstraction, then only configurations of devices adjacent to the change ever need to be updated.

@@

\sysname takes as input the network's abstract topology and its policy. The policy has two components: the routing policy describes how traffic should flow, and the fault-tolerance policy describes how many simultaneous failures the network can withstand without loosing connectivity.

\para{Abstract topology}
Abstract topologies in \sysname capture connectivity invariants of the network that hold both when network is first instantiated and as it evolves. These invariantes are encoded in the form of a graph homomorphism annotated with logical constraints about node and edge multiplicities. 
%
We designed these abstractions to be able to precisely capture most real network topologies, while being amenable to the fault tolerance analysis and configuration synthesis that is done by our compiler.

@@

We introduce our topology abstractions using the networks from \S\ref{sec:motivation}. Topology abstractions consist of several components.
The first is a role-based abstraction that allows an operator to map routers in the concrete network to roles in the abstract network. Figure~\ref{fig:example3} shows an example of an abstraction for both datacenter networks from \S\ref{sec:motivation}. In particular the concrete networks are abstracted into a new topology with 5 different roles: local ToR (TL), global ToR (TG), local aggregator (AL), global aggregator (AG), and spine (S).

More specifically, a network topology is a graph $G$ = ($V, E$), which consists of a set of vertices $V$ and a set of directed edges $E \colon V \times V$. A role-based abstraction is a graph homomorphism from $G$ to an abstract graph $G^A$ = ($V^A$,$E^A$). A graph homorphism $f : G \rightarrow G^A$ maps each node in the concrete graph to a node in the abstract graph such that, whenever $(u,v) \in E$, then $(f(u),f(v)) \in E^A$.

@@

The role-based abstraction therefore overapproximates the connectivity of the underlying concrete graphs. On its own, this abstraction loses a lot of information about the underlying structure of the concrete network, making it difficult to reason precisely about fault-tolerance.
%which would hinder a useful analysis of its fault-tolerance. 
For example, with this abstraction any spine router in a concrete network may or may not be connected to any aggregator router.
%
In order to refine the abstraction to capture more precisely the kinds of concrete networks of interest,
we introduce several additional concepts. The first is the idea of topology hierarchy captured by P and Q, which say that nodes in the tor and aggregator roles are grouped into a number of smaller pods.
The second is a notion of node and edge \emph{multiplicities}. Each edge (and node) is labelled
with a symbolic variable (\EG, $e1, e2, \ldots$) that denotes a constraint on the number of edges (and nodes) that may appear in any given concrete network. Operators can capture concrete network invariants by adding constraints on the symbolic variables using logical formula.

For example, in Figure~\ref{fig:example3}, the first two constraints: $e1 = AG$ and $e2 = TG$ say that, within any pod P, the number of outgoing edges from a node in the $TG$ role to a node in the $AG$ role is equal to the number of nodes in the $AG$. Similarly, the number of outgoing edges from $AG$ to $TG$ is equal to the number of nodes in the $TG$ role. These constraints capture the fact that, within any given pod, the global aggregators and ToRs are in a full mesh. Furthermore, the constraints $AG = AL$ and $AG \leq S$ ensure that, within pods P and Q, the $AG$ and $TG$ roles have the same number of routers, which is less than the number of routers in the spine role $S$.
%
The constraint $e3 \geq 2$ says that, in each pod, each aggregator node has at least 2 outgoing edges to nodes in the spine role. Symmetrically, the constraint $e4 \geq 1$ says that, for each pod P, each node in the spine role has at least one outgoing edge to a node in the $AG$ role. Similar constraints appear for the local aggregator role. The constraint $2 \leq S \leq 4$ makes explicit the possibility for growth, for example by growing the network from that of Figure~\ref{fig:example} to that of Figure~\ref{fig:example2}. In general, we need not bound the number of spine routers to admit more concrete topologies at the expense of analysis precision. Finally, we include the constraints $(S \mod AG) = 0$, and $(S \mod AL) = 0$ simply to show that constraints do not have to inequalities. In general operators can write constraints using logical formula for any theory supported by modern SMT solvers.

\ryan{mention mincut}

@@

\para{Routing policy} Routing policies in \sysname consist of an ordered sequence of both a predicate matching various classes of traffic and a list of ranked constraints on the paths that the traffic matching the predicate should prefer to take through the network. We borrow syntax from Propane~\cite{propane} but, instead of concrete prefixes and nodes, \sysname specifications can use abstract prefixes and nodes.

Let us see how to configure each of the datacenter networks from \S\ref{sec:motivation} by writing a \sysname routing policy over the abstract topology. We can capture the basic routing behavior for the datacenters with the following policy:
%
\begin{code}
\Define Routing =
    \$GP  \Path \End(TG)
    \$LP  \Path \End(TL)
    \True \Path \End(\Out) & \Exit(P1 \Prefer P2)
\end{code}
\noindent%

The first line introduces a prefix \emph{template} variable. It says that traffic for each global prefix associated with the template variable $\CD{\$GP}$ has the constraint that it should follow a path that ends at its corresponding destination router contained in the $\CD{TG}$ role. The second line does the same for all local prefixes.
%
The final rule matches all other IP prefix destinations and allows traffic to follow a path that leaves the datacenter, ending up at some outside location, through either of its peers $\CD{P1}$ or $\CD{P2}$ with a preference for leaving through $\CD{P1}$ over $\CD{P2}$. The $\CD{\Prefer}$~symbol indicates that traffic should satisfy the constraint on the left whenever possible (\IE, a path exists in the network), and only resort to the backup (right) when this is not possible due to network failures.%

Next we can capture the constraint that traffic for local prefixes must stay local to the datacenter.
This is written as the following \sysname constraint:
%
\begin{code}
\Define Local =
    \$LP \Path \Always(\In)
\end{code}
\noindent%
%
The Local constraint applies to any traffic destined for a local prefix described by the \$LP prefix template variable and adds the constraint that the traffic must follow a path that matches an internal location (\IE, a router inside the datacenter) at each hop of the path. This prevents traffic from ever entering from or leaving through an external peer.

It is also common practice to prevent peers (\EG, $\CD{P1}$ and $\CD{P2}$) from communicating by sending data through the datacenter. Such a constraint can be added to \sysname in the following way:
%
\begin{code}
\Define Peer = \{P1,P2\}
\Define BlockPeers =
    \True \Path !(\Enter(Peer) & \Exit(Peer))
\end{code}
\noindent%
%
We define a $\CD{Peer}$ to be either $\CD{P1}$ or $\CD{P2}$ and disallow paths where traffic both enters the datacenter through a peer and exits through a peer. Finally, we can combine these constraints together as follows:

\begin{code}
Routing & Local & BlockPeers & \Agg(PG, \In -> \Out)
\end{code}
\noindent%

This tells \sysname to find BGP router configurations that satisfy the conjunction of the Routing and Local constraints. It also declares an additional constraint that we want to perform prefix summarization (as PG) for global prefixes at the border of the datacenter (\IE, along any edge starting inside our datacenter and ending at a peer outside the datacenter).

@@

%For example, the combining the Routing and Local constraints in the data center policy from \S\ref{sec:overview} would result in the following \sysname policy:
%
%\begin{code}
%\$PG  \Path \End(TG),
%\$PL  \Path \End(TL) \ensuremath{\cap} \Always(\In),
%\True \Path \Exit(Peer1) \Prefer \Exit(Peer2)
%\end{code}
%\noindent%
%

@@

%By combining the tagging, filtering and preference mechanisms, we can generate safe BGP configurations. In general, configurations are generated by combining the per-prefix configurations in the same order specified in the policy.
Figure~\ref{fig:bgp-configs} shows part of the final configuration for spine routers for the concrete and abstract policies.
%
For prefix $0.0.0.0/[0..32]$, the spine routers will match advertisements from peer $P1$ and $P2$. The match for $P1$ is preferred since it has a higher BGP local-preference attribute (110). If an advertisement from $P1$ is available the spine uses this and attaches the community tag $(1,0)$ and advertises the route to all its peers. If only an advertisement from the backup $P2$ is available, then it attaches the tag $(0,1)$ and advertises this to all peers instead.
%
There is also additional policy for the Product Graphs from constraints for the local and global prefixes. In the template configuration, it matches any global prefix from any internal peer and re-advertise the route to all of its peers. For any local prefix it will allow an advertisement from any internal peer, and re-advertise the route to only other internal peers (thus implementing the Local constraint). The concrete configurations have a similar configuration for each local and global prefix.
